---
title: "Data Viz FINAL!"
subtitle: "Data Visualization (STAT 302)"
author: "Anna Wagman"
output:
  html_document:
    toc: true
    toc_float: true
    highlight: "tango"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction:
As an avid Netflix consumer and lover of entertainment, I examined data about the titles available of Netflix. Throughout these visualizations, I explore the relative amounts of Movies vs TV Shows, and then compare specific variables between the two such as "listed_in" and "duration". I created two distinct visualizations displaying the streaming availability around the world. I extract specific variables such as "date_added" and "country" to explore metrics and changes over time. Throughout my seven explorations, I will highlight key data and statistics about the content on Netflix**
<br>
<br>
**Data source: <https://www.kaggle.com/shivamb/netflix-shows>**



```{r, message=FALSE, include = FALSE, warning=FALSE}
#Load packages:

library(tidyverse)
library(janitor)
library(skimr)
library(ggplot2)
library(RColorBrewer)
library(lubridate)
library(dplyr)
library(ggmap)
library(viridis)
library(cowplot)

#import data:
netflix <- read_csv("data/netflix_titles.csv")

#Make into tibble to look at data
head(netflix)
ncol(netflix)
nrow(netflix)
n_distinct(netflix)
colnames(netflix)
str(netflix)
```

# Visualization 1: Total Movies & TV Shows per Country
### Q1: What Countries have the most content? Is there a large disparity? 

```{r, message = FALSE, warning=FALSE, include = FALSE}
#add n = amount of TV + Movies in each country
#create subset of 10 countries with highest n, most titles
#graph in descending order with graphics (color, size, labels)


countries_plot <- netflix %>%
  filter (!str_detect (country, ",")) %>% #multiple countries seperated by commas
  group_by (country) %>% #group by country
  count() %>% #add total in each group = country
  arrange(desc(n)) %>% #descending order (decreasing)
  head(20) %>% #select n for highest 20
  
  ggplot() + 
  geom_col(fill = "#C60B0B", color = "black", #color 
    aes(y = reorder(country, n), x = n)) + #order highest -> lowest
  labs( 
    title = "Movies & TV Shows per Country", #labels
    x = "Number of Titles",
    y = "Country"
  ) +
  geom_label(aes(y = reorder(country,n), x = n, label = n), size = 2.3) + #label n on each bar
  theme_classic() 

```

```{r, message = FALSE, warning=FALSE}
countries_plot
```

## Plot 1b. World Map of Available content

```{r, message=FALSE, include=FALSE, warning=FALSE}
#calculate number per country
netflix_map <- netflix %>% 
  filter(!(is.na(country) | country == "")) %>% 
  group_by(country) %>% 
  summarise(number_of_titles = length(unique(title)), .groups = "drop") 

#map world
worldmap_plot <- map_data("world")

#change country names to match
netflix_map$country <- recode(netflix_map$country, 
                             "United States" = "USA",
                             "United Kingdom" = "UK")
  
#join netflix data and world map data
netflix_map <- full_join(netflix_map, worldmap_plot, by = c("country" = "region"))

#make plot!
netflix_world_map <- ggplot(
  data = netflix_map,
  aes(a = country, b = number_of_titles)
  ) +
  geom_polygon(
    aes(x = long, y = lat,
      group = group, fill = number_of_titles
    )
  ) +
  scale_fill_viridis(discrete = FALSE) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank()
    ) +
  ylim(-50, 90) +
  theme_minimal()
```


```{r, message = FALSE, warning = FALSE}
netflix_world_map
```


# Visualization 2: Movies vs TV shows 

### Q2. Does Netflix have more Movies or TV Shows? What is the difference?


```{r, message = FALSE, warning=FALSE, include=FALSE}
#count number of movies and TV shows 
#create Pie chart to compare the two 

Amount_MvT <- netflix %>%
group_by(type) %>% #group by Movie and TV Show 
summarize(number_of_each = n()) #add up each
```

```{r, message = FALSE, warning=FALSE}
Amount_MvT
```

```{r, message = FALSE, warning=FALSE, include = FALSE}
MvT_plot <- ggplot(Amount_MvT, aes(x = "", y = number_of_each, fill = type)) +
  geom_bar(stat = "identity") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = number_of_each), 
            position = position_stack(vjust = .5)) +
  labs(
    title = "Netflix: Amount of Movies vs TV shows",
    y = "Number of each",
    x = NULL
  ) +
  scale_fill_manual (values = c("#61DCDE", "#F57CF4")) +
  theme_minimal()
```

```{r, message = FALSE, warning=FALSE}
MvT_plot
```


# Visualization 3: Amount of titles over time 

### Q3. How many titles get added each year? Has it been increasing or decreasing over time?

```{r, message = FALSE, warning=FALSE, include = FALSE}
##read dates
netflix$date_added <- as.Date(netflix$date_added,format = '%B %d, %Y')

#Find amount of TV shows and Movies added each year
#Plot amount of (Movies & TV shows)  by (time) with a scatter plot
mvt_added <- netflix %>%
group_by(type,date_added)%>%
summarise(number_released=n())%>%
mutate(total_shows=cumsum(number_released))
head(mvt_added)
```


```{r, message = FALSE, warning=FALSE, include = FALSE}
added_plot <- ggplot(mvt_added,aes(date_added,total_shows,color = type)) +
geom_line() +
scale_x_date(breaks = '1 years',date_labels = '%Y',limits = as.Date(c("2007-01-01","2023-01-01"))) +
theme(plot.title = element_text(size = 5),
      axis.text.x = element_text(size = .5),
      axis.text.y = element_text(size = 5),
      axis.title = element_text(size = 5),
      legend.text=element_text(size = 1),
      legend.title=element_text(size = 5)
      ) +
  labs(
    x = "year",
    y ="Amount of titles added",
    title="Amount of Movies vs TV Added per year"
    ) +
  theme_minimal()
```
```{r, warning=FALSE, message=FALSE}
added_plot
```


# Visualization 4: Genres of Movies and TV Shows 

### Q4. What are the most common genres of Netflix? Are the Genres different for Movies vs TV Shows? 

```{r, message = FALSE, warning=FALSE, include = FALSE}
#select genre = listed_in 
#count how many in each 
#select 10-15 most common Genres to plot 

MvT_Genres <- netflix %>%
  select(listed_in) %>%
  mutate(listed_in = str_split(listed_in,','))%>%
  unnest(listed_in) %>%
  group_by(listed_in) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(15) %>%
  
  ggplot() + 
  geom_col(aes(y = reorder(listed_in,n), x = n, fill = listed_in)) +
  labs(title = "Genres of Movies and TV Shows",
     x = "Amount of Titles",
     y = "Genre") +
  geom_label(aes(y = reorder(listed_in, n), x = n, label = n), size = 2.2) +
  theme_minimal() 

```
```{r, message = FALSE, warning=FALSE}
MvT_Genres
```


## Plot 4b: Genres of Movies vs TV Shows (compared)

```{r, message = FALSE, warning=FALSE, include = FALSE}
#filter by type = "Movie" and select genre = "listed_in"
#count number of each genre 
#plot number of each genre in descending order 

#select movies and group by genre
Movie_Genres <- netflix %>%
  filter(type == "Movie") %>%
  mutate(listed_in = str_split(listed_in,','))%>%
  unnest(listed_in) %>%
  group_by(listed_in) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(18) %>%
  
#Plot amount of each genre of Movie
  ggplot() + 
  geom_col(fill = "lightblue", color = "black",
    aes(y = reorder(listed_in, n), x = n)) +
  labs(
    title = "Movie Genres",
    x = "Amount of Movies",
    y = "Genre"
  ) +
  geom_label(aes(y = reorder(listed_in, n), x = n, label = n), size = 2.2) +
  theme_minimal()
  
```

```{r, message = FALSE, warning=FALSE, include=FALSE}
#filter by type = "tv show"
#select genre = listed_in, add up n 

TV_Genres <- netflix %>%
  filter(type == "TV Show") %>%
  mutate(listed_in = str_split(listed_in,','))%>%
  unnest(listed_in) %>%
  group_by(listed_in) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(18) %>%
  
#graph amount each genre of TV Show
  ggplot() + 
  geom_col(fill = "lightpink", color = "black",
    aes(y = reorder(listed_in, n), x = n)) +
  labs(
    title = "TV Show Genres",
    x = "Amount of Shows",
    y = "Genre"
  ) +
  geom_label(aes(y = reorder(listed_in, n), x = n, label = n), size = 2.2) +
  theme_minimal()


```

```{r, message = FALSE, warning=FALSE}
Movie_Genres
TV_Genres
```



# Visualization 5: Release Date vs Date Added
### Q5. How long after something is released is it added to Netflix? Is it usually within the year?

```{r, message = FALSE, warning=FALSE, include=FALSE}
#mutate data to subtract the dates
netflix_dates <- read_csv("data/netflix_titles.csv") %>%
  mutate(date_added = mdy(date_added)) %>% clean_names() %>%
  mutate(year_diff = year(date_added) - release_year)


netflix_dates_tibble <- netflix_dates %>% 
  count(year_diff, sort = T)

as.tibble(netflix_dates_tibble)


```


```{r, message = FALSE, warning=FALSE, include=FALSE}
library(RColorBrewer)

#mutate data to subtract the dates
netflix_dates2 <- read_csv("data/netflix_titles.csv") %>%
  mutate(date_added = mdy(date_added)) %>% clean_names() %>%
  mutate(year_diff = year(date_added) - release_year) %>%
  count(year_diff) %>%
  arrange(desc(n)) %>%
  head(25) %>%
  
  ggplot() +
  geom_col(
    aes(fill = year_diff, x = reorder(year_diff, n), y = n)) +
  labs(
    title = "Release Date vs Year Added",
    x = "Difference in years (added-release)",
    y = "Amount of titles"
  ) +
  theme_minimal()

 



```

```{r, message = FALSE, warning=FALSE}
netflix_dates2 
```

# Visualization 6: Notable Directors & Actresses/Actors

### Q6. Who are the most frequence Directors on Netflix? With how many films? Who are the most frequent Actresses and Actors?
```{r, message = FALSE, warning=FALSE, include=FALSE}
directors <- netflix %>%
  filter(!str_detect (director, ", ")) %>%
  group_by(director) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(15) %>%
  
  ggplot() +
  geom_col(fill = "#78DB79", color = "black",
           aes(y = reorder(director, n), x = n)) +
  labs(
    title = "Directors",
    x = "Number of titles",
    y = "Director"
  ) +
  theme_minimal() +
  geom_label(aes(y = reorder(director, n), x = n, label = n), size = 2.2)
```

```{r, message = FALSE, warning=FALSE, include=FALSE}
#most common actors and actresses
actors <- netflix %>%
  mutate(cast = str_split(cast, ", ")) %>%
  unnest(cast) %>%
  group_by(cast) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(15)

actors_removeNA <- actors[-1, ] 
  
  
actors_plot <- ggplot(actors_removeNA) +
  geom_col(fill = "#B080DB", color = "black",
           aes(y = reorder(cast, n), x = n)) +
  labs(
    title = "Actresses & Actors",
    x = "Number of titles",
    y = "Actress / Actor"
    ) +
  theme_minimal() +
  geom_label(aes(y = reorder(cast, n), x = n, label = n), size = 2.2) 
```

```{r, message = FALSE, warning=FALSE}
plot_grid(actors_plot, directors) 
```

# Visualization 7. TV Shows with the most seasons
### Q7. Which TV shows on netflix have the most seasons? How many seasons are there of that show?

```{r, message = FALSE, include = FALSE, warning = FALSE}
netflix_seasons <- netflix %>%
  filter(type == "TV Show") %>%
  select(title, duration) %>%
  mutate(duration = extract_numeric(duration)) %>%
  arrange(desc(duration)) %>%
  head(15) %>%
  
  ggplot() +
  geom_col(fill = "#F9F054", color = "red",
    aes(y = reorder(title, duration), x = duration)) +
  labs(
    title = "Longest Running TV Shows",
    x = "Amount of Seasons",
    y = "TV Show"
  ) +
  geom_label(aes(y = reorder(title, duration), x = duration, label = duration), size = 2.2) +
  theme_minimal() 
```

```{r, warning = FALSE, message = FALSE}
netflix_seasons
```


